ECCCos from the Black Box:\\ Faithful Explanations through\\ Energy-Constrained Conformal Counterfactuals

### 2023-09-26 paper/appendix.tex
mostly done
nearly there now

\begin{equation}\label{eq:biased-sgld}
  \begin{aligned}
    \hat{\mathbf{x}}_{j+1} &\leftarrow \hat{\mathbf{x}}_j - \frac{\epsilon}{2} \mathcal{E}_{\theta}(\hat{\mathbf{x}}_j|\mathbf{y}^+) + \sigma \mathbf{r}_j, && j=1,...,J
  \end{aligned}
\end{equation}

Consistent with~\citet{grathwohl2020your}, we have specified $\epsilon=2$ and $\sigma=0.01$ as the default values for all of our experiments. The number of total SGLD steps $J$ varies by dataset (Table~\ref{tab:ebmparams}). Following best practices, we initialize $\mathbf{x}_0$ randomly in 5\% of all cases and sample from a buffer in all other cases. The buffer itself is randomly initialised and gradually grows to a maximum of 10,000 samples during training as $\hat{\mathbf{x}}_{J}$ is stored in each epoch~\citep{du2019implicit,grathwohl2020your}.

@@

\subsubsection{Compute}

To enable others to easily replicate our experiments, we have chosen to work with small neural network architectures and randomly undersampled the \textit{MNIST} dataset (maintaining class balance). All of our experiments could then be run locally on a personal machine. The longest runtimes we experienced for model training and counterfactual benchmarking were on the order of 8-12 hours (\textit{MNIST} data). For the synthetic data, all experiments could be completed in less than an hour.

We have summarised our system information below:

@@

@@

Table~\ref{tab:results-full} reports all of the evaluation metrics we have computed. Table~\ref{tab:results-full-valid} reports the same metrics for the subset of valid counterfactuals. The `Unfaithfulness' and `Implausibility' metrics have been discussed extensively in the body of the paper. The `Cost' metric relates to the distance between the factual and the counterfactual. The `Redundancy' metric measures sparsity in is defined as the percentage of features that remain unperturbed (higher is better). The `Uncertainty' metric is just the average value of the smooth set size penalty (Equation~\ref{eq:setsize}). Finally, `Validity' is the percentage of valid counterfactuals.

\import{contents/}{table_all.tex}

\import{contents/}{table_all_valid.tex}

>>>>

\begin{equation}\label{eq:biased-sgld}
  \begin{aligned}
    \hat{\mathbf{x}}_{j+1} &\leftarrow \hat{\mathbf{x}}_j - \frac{\phi}{2} \mathcal{E}_{\theta}(\hat{\mathbf{x}}_j|\mathbf{y}^+) + \sigma \mathbf{r}_j, && j=1,...,J
  \end{aligned}
\end{equation}

Consistent with~\citet{grathwohl2020your}, we have specified $\phi=2$ and $\sigma=0.01$ as the default values for all of our experiments. Here we have deliberately departed slightly from the notation in Equation~\ref{eq:sgld} to emphasize that we use fixed values for $\phi$ and $\sigma$, consistent with the related literature. The number of total SGLD steps $J$ varies by dataset (Table~\ref{tab:ebmparams}). Following best practices, we initialize $\mathbf{x}_0$ randomly in 5\% of all cases and sample from a buffer in all other cases. The buffer itself is randomly initialised and gradually grows to a maximum of 10,000 samples during training as $\hat{\mathbf{x}}_{J}$ is stored in each epoch~\citep{du2019implicit,grathwohl2020your}.

@@

\subsection{Compute}

To enable others to easily replicate our experiments, we have chosen to work with small neural network architectures and randomly undersampled the \textit{MNIST} dataset (maintaining class balance). All of our final benchmarks could then be run locally on a personal machine, but grid searches and the sheer number of datasets required us to move to high-performance computing clusters to conduct our experiments efficiently.

\subsubsection{Local runs}

The longest runtimes we experienced for model training and counterfactual benchmarking for a single benchmark were on the order of 12-24 hours (\textit{MNIST} data). For the synthetic data, single benchmarks could be completed in less than an hour. We have summarised our system information below:

@@

\subsubsection{HPC}

We used two large clusters, which we do not cite or mention by name at this point to not interfer with the double-blind review process. All of our grid searches were multi-processed on 100 CPUs at 4-8GB or memory each.

@@

Tables~\ref{tab:results-linearly-separable} to~\ref{tab:results-fashion-mnist} reports all of the evaluation metrics we have computed. Tables~\ref{tab:results-linearly-separable-valid} to~\ref{tab:results-fashion-mnist-valid} reports the same metrics for the subset of valid counterfactuals. The `Unfaithfulness' and `Implausibility' metrics have been discussed extensively in the body of the paper. The `Cost' metric relates to the distance between the factual and the counterfactual and is measured using the L1 Norm. The `Redundancy' metric measures sparsity in is defined as the percentage of features that remain unperturbed (higher is better). The `Uncertainty' metric is just the average value of the smooth set size penalty (Equation~\ref{eq:setsize}). Finally, `Validity' is the percentage of valid counterfactuals.

\import{contents/}{table-linearly-separable.tex}

\import{contents/}{table-circles.tex}

\import{contents/}{table-moons.tex}

\import{contents/}{table-california-housing.tex}

\import{contents/}{table-gmsc.tex}

\import{contents/}{table-german-credit.tex}

\import{contents/}{table-mnist.tex}

\import{contents/}{table-fashion-mnist.tex}

\import{contents/}{table-linearly-separable-valid.tex}

\import{contents/}{table-circles-valid.tex}

\import{contents/}{table-moons-valid.tex}

\import{contents/}{table-california-housing-valid.tex}

\import{contents/}{table-gmsc-valid.tex}

\import{contents/}{table-german-credit-valid.tex}

\import{contents/}{table-mnist-valid.tex}

\import{contents/}{table-fashion-mnist-valid.tex}
